Concave reduction in Self-Healing Swarms

### 2018-11-14 paper.tex
much tidying of mathematics

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{relsize}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{array}
%\usepackage{caption}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage{fixltx2e}
\usepackage{float}
\usepackage{url}
%\usepackage{becs}
\usepackage{listings}
\lstset{%
  basicstyle=\normalsize,
  numbers=left,
  numberstyle=\normalsize,
  stepnumber=1,
  xleftmargin=2em
}
\usepackage{color}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\let\latexvec=\vec
\let\vec=\latexvec
\interdisplaylinepenalty=2500
\newtheorem{definition}{Definition}
\newtheorem{exampleb}{Example}
\newcommand{\subsubsubsection}[1]{\medskip\par\textit{#1:}}
\newcommand{\card}[1]{\vert{#1}\vert}
\newcommand{\magn}[1]{\Vert{#1}\Vert}
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Concave reduction in Self-Healing Swarms}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, Michael Brockway\authorrefmark{1}, Alun Moon\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, Newcastle upon Tyne, NE1 8ST}
\markboth
{Eliot \headeretal: Concave reduction in Self-Healing Swarms}
{Eliot \headeretal: Concave reduction in Self-Healing Swarms}
\corresp{Corresponding author: Neil Eliot (e-mail: neil.eliot@northumbria.ac.uk).}
\begin{abstract}
%\boldmath
Swarms consist of many agents that interact using a simple set of rules that create emergent behaviours. One such behaviour makes a swarm tolerant of failures. Failures occur for many reasons, for example; resource exhaustion, component failure, or a disruption from an external event. The loss of agents reduces the size of a swarm and may create an irregular structure. A swarm's structure can also be irregular due to initial conditions, loss of agents or due to an obstacle. These changes in the structure or size do not stop a swarm from functioning but can adversely affect the efficacy. This paper introduces and demonstrates emergent behaviour that can counter the effect of agent loss or irregular structures. The behaviour causes concave deformities to be removed. This technique is implemented by each agent identifying its position in relation to its neighbours. If the agent meets a defined criterion, based on the local geometry of the agent's neighbours to decide when an anomaly is present, the agent's `normal' movement is altered in such a way as to remove a gap between the neighbours. These changes in direction of agents cause the swarm, as a whole, to be reorganised into a more uniform structure with no internal voids.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.
% Note that keywords are not normally used for peer review papers.
\begin{IEEEkeywords}
Swarming, Swarm Dynamics, Mobile Sensor Networks, Algorithms.
\end{IEEEkeywords}
\titlepgskip=-15pt
\maketitle
\section{Introduction}\label{sec:ConcaveReduction}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively.
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two \textit{field effects} to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. \textit{Field effects} have ranges around an agent to determine the effect other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}) the effects are derived from these ranges as vectors. All ranges must fall within the sensing capabilities of the agent ($S_b$). It is usual for the cohesion field to have a radius ($C_b$) which is larger than the repulsion radius ($R_b$). When an agent ($b'$) moves into the \textit{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has its repulsion vector applied and the agent will tend towards moving away from the obstacle.
\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}
When cohesion and repulsion (\textit{inter-agent vectors}) are the only field effects creating a swarming effect the number of stable structures that can develop is limited. These structures, which are constructed from triangles, are either straight edges or partial lattices. Partial lattices can create concave anomalies (\textit{dents}) and convex anomalies (\textit{peaks}) in a perimeter~(Fig.~\ref{concave:SwarmStableShape1}). An anomaly is any construct within the swarm that is not a hexagonal lattice or an agent distribution that causes the swarm to deviate from having only convex or flat edges. The advantage of creating a well structured swarm can be realised in two classes of swarm application; \textit{Reconnaissance}, by creating a more uniform structure that eliminates blank spots, and \textit{Object Containment}, when using a swarm to surround a target.

@@

\section{Swarm Model}\label{sec:Swarm Model}
To describe the removal of voids a swarming model needs to be defined. Equation~\ref{eq:BotPhysics1} defines a weighted model that includes cohesion, repulsion, direction and obstacle avoidance ($v_c(b), v_r(b), v_d(b), v_o(b)$). The weightings $k_c, k_r, k_d, k_o$ allow each component to be scaled to tailor the swarming effect.

\begin{equation}\label{eq:BotPhysics1}
  v(b) = k_cv_c(b) + k_rv_r(b) + k_dv_d(b) + k_ov_o(b)
\end{equation}

\textit{Repulsion} $v_{r}(b)$, defined in Equation~\ref{eq:Repulsion1}, is the
directional movement required to prevent agents colliding. $\mathcal{R}_b$ is
the set of agents that are within an agent's $b$ repulsion range.

\begin{equation}\label{eq:Repulsion1}
v_{r}(b) = 
\frac{1}{|\mathcal{R}_b|}
\left(
	\mathlarger{\mathlarger{\sum_{b' \in \mathcal{R}_b}}}
{\left( 1-\frac{|bb'|}{R_b} \right)}
{bb'}
\right)
\end{equation}

\textit{Cohesion} $v_{c}(b)$, defined in Equation~\ref{eq:FlyToCentre1},
calculates the movement that makes an agent move towards other agents to form
a cohesive structure. $\mathcal{C}_b$ is the set of agents that are within an agent's ($b$) cohesion range.

\begin{equation}\label{eq:FlyToCentre1}
	v_{c}(b) = \frac{1}{|\mathcal{C}_b|}\left({\mathlarger{\sum_{b' \in
	\mathcal{C}_b}}{bb'}}\right)
\end{equation}

\textit{Direction} $v_d(b)$, defined in Equation~\ref{eq:Direction}, generates a directional vector for an agent to move towards a destination defined as $d$.

\begin{equation}
\label{eq:Direction}
v_d(b) = bd
\end{equation}

@@

Equation~\ref{eq:Obstacle2} shows the repulsion vector $v_o(b)$ for an agent.
$\mathcal{O}_b$ is the set of obstacles that are within the range of agent $b$.
The obstacles are identified using the distance between an agent and an
obstacle $\magn{ob}$ and comparing the result to the fixed obstacle repulsion
field~$O_b$, so $\forall o \in \mathcal{O}_b : \magn{ob}\leq O_b$.
The applied repulsion is calculated by scaling the normalised sum of the normalised vectors $\widehat{ob}$ by $O_b$. Note that $\string^$ is the equivalent of $\hat{v} = \frac{v}{\magn{v}}$ the normalised vector.

@@

\begin{eqnarray}\label{eq:Obstacle2}
  v_o(b) & = & O_b \widehat{d_o} \\
	\mathrm{where~}  d_o & = & \sum_{o\in \mathcal{O}_b } \widehat{ob}
  \nonumber
\end{eqnarray}

An agent's \textit{movement vector} is the sum of all the component vectors as shown in Equation~\ref{eq:BotPhysics1}. This is a similar technique as used by Hashimoto et. al.~\cite{HAY:08}. For a vector to be used for movement it be normalised before the agent's speed can be applied. The resultant \textit{movement vector} $b_{pos}$, as defined in Equation~\ref{eq:AgentMovement}, is calculated using unit time, speed and the normalised \textit{movement vector}.

\begin{equation}\label{eq:AgentMovement}
  b_{pos}=s_{b}t\widehat{v(b)} 
\end{equation}

@@

The detection mechanism detects both the outer edge of a swarm and any internal features~(voids) that satisfy the same set of conditions. It is therefore possible to have both voids and islands of agents within the same swarm~(Fig.~\ref{fig:PerimeterBots2}). Voids are best defined as perimeters that are concave in nature and inside another perimeter. McLurkin describes two types of perimeters convex and concave~\cite{MD:09}. A convex perimeter is an edge where the average angle of the agent's exposed faces are~$> 180^\circ$. A concave perimeter is where the average exposed angle is~$< 180^\circ$.

@@

An agent $b$ in the swarm has a sensor that detects other agents in the swarm
determining their range $r$  and direction $\beta$.  From this data the agent generates a
set of neighbours that are within a specific range.  Usually the range of the
cohesion field $C_b$.

The Agent creates a set of neighbours containing range and bearing pairs
$(\beta,r)$ as shown in equation
\ref{set:Nb}
\begin{equation}
\mathcal{N} = \{ (r,\beta) \ldots \}
\label{set:Nb}
\end{equation}
This set is sorted into a sequence $\mathcal{P}_A$ with the bearing values
increasing
\begin{equation}
	\mathcal{P}_a = \langle (r_0,\beta_0),\ldots,(r_n,\beta_n) \rangle
\end{equation}
such that $\forall i \in \{0\ldots n-1\}: \beta_i<\beta_{i+1}$

This set of agents forms the perimeter of an enclosing polygon of agent $b$. 
Each consecutive pair of agents (in the sequence) defines an edge, which has
length $d$ and an angle $\theta$ given by the difference in bearings of
succesive neighbours.  The sequence of edges
that forms this polygon is
\begin{equation}
	\mathcal{P}_e = \langle (d_0,\theta_0), \ldots , (d_n,\theta_n) \rangle
\end{equation}
where
\begin{equation}
	\theta_i = \beta_{i+1} - \beta_i
\end{equation}
the index addition is modulo $|\mathcal{N}|$ making $\beta_0$ the successor
bearing to $\beta_n$ ($n+1 = 0$).  The angles $\theta$ must lie in the range
$0<\theta\leq2\pi$.
This restriction on the values of $\theta$ enforce a condition that
\begin{equation}
	\sum\theta_i = 2\pi
\end{equation}
The length of a perimeter edge is given by the cosine rule
\begin{equation}
	d_i^2 = r_{i+1}^2 + r_i^2 -2r_{i+1}r_i \cos\theta_i
\end{equation}

An agent is on the perimeter of the swarm if it is not enclosed by the polygon
defined in $\mathcal{P}_e$.  Simple geometry shows that this is the case given
by the predicate \ref{enclosing}
\begin{equation}
	\exists \theta_i \in \mathcal{P}_e : \theta_i\geq\pi
	\label{enclosing}
\end{equation}
The polygon is considered `open' if two succesive agents on the perimeter 
cannot see each onther, their separation $d$ is greater that the range of the
attractive field.  An open polygon does not enclose the agent $b$ so it
considered to be on the perimeter.

The agent $b$ is on the perimeter of the swarm if the predicate in equation
\ref{perimeter} is true.
\begin{equation}
	\exists d_i\in\mathcal{P}_e:d_i>d_a \vee
	\exists\theta\-i\in\mathcal{P}_e:\theta_i\geq\pi
	\label{perimeter}
\end{equation}
An agent is at the apex of a concave region of the perimeter if
\begin{equation}
	\exists(\theta_i,d_i)\in\mathcal{P}_e : d_i>d_a\wedge\theta_i<\pi
\end{equation}

\textit{Orientation independent}  If the agent $b$ is rotated through an angle of $\gamma$ then 
the bearings are rotated by $-\gamma$, \[ \beta_i\mapsto\beta_i-\gamma \]
The  angle between successive agents is now
\[
	\theta_i  =  (\beta_{i+1}-\gamma) - (\beta_i-\gamma)
	 = \beta_{i+1}-\beta_i-\gamma+\gamma
	 = \beta_{i+1}-\beta_i
\]

@@

\begin{figure}
\begin{algorithmic}[1]
%\tiny
\Procedure{CheckVisibility}{$b,angles$}
\State $b.gap \leftarrow \emptyset$
\State $b.concave = False$
\For{$i \leftarrow 0~\texttt{To}~size(angles)-1$}
\If{i == size(angles)-1}
\If{$cosrule(b,angles[size(angles)-1][0],angles[i][0])$}
\If{$angle(b,angles[size(angles)-1][0],angles[i][0]) < 180)$}
\State $b.concave = True$
\State $b.gap = (angles[size(angles)-1][1],angles[i][1])$
\EndIf
   \State\textbf{return} $True$
   \EndIf
\EndIf
\If{$cosrule(b,angles[i + 1][0],angles[i][0])$}
  \If{$angle(b,angles[i + 1][0],angles[i][0]) < 180$}
\State $b.concave = True$\;
\State $b.gap = (angles[i + 1][1],angles[i][1])$
  \EndIf
  \State\textbf{return} $True$
\EndIf
\EndFor
\State\textbf{return} $False$
\EndProcedure
\end{algorithmic}
\caption{Check Visibility}
\label{algo:checkVisibility2}
\end{figure}

@@

When an agent is identified as being a component of a concave characteristic~(Fig.~\ref{algo:checkVisibility2}) the normal \textit{movement-direction vector} is replaced by a \textit{concave reduction vector}. This new vector causes the agent to move in a direction that will reduce or remove a concave edge by moving the agent towards the identified gap. The effect of this will be to either straighten an outer perimeter or reduce/remove a void. This change in direction effects the distance and magnitude variances due to the changes induced. Figures~\ref{fig:InterAgentEffect1} and \ref{fig:InterAgentEffect2} shows this effect in more detail. Figure~\ref{fig:InterAgentEffect1}~shows the initial positions of the agents before the concave reduction is applied to the agent. Figure~\ref{fig:InterAgentEffect2}~shows the effects on the relationship with the agent's neighbours. The aggregate change is an increase in the inter-agent distances and an increase in the resultant magnitude effects.

@@

\section{Concave reduction mathematical model}\label{concave:ConcaveVoidReduction1}
As part of identifying a perimeter each agent's neighbours must be sorted by angle. A pair of agents are identified as a set $b.gap$ ($b.gap \equiv G_b$). Equation~\ref{eq:ConcaveVoidPhysics3} calculates the centroid of the gap agents.

Equation~\ref{eq:ConcaveVoidPhysics1} generates the dictionary set of all the neighbours along with an angle that each of the agent's neighbours make with the first detected neighbour. $nbr(b)$ returns a set consisting of an agent's ($b$) neighbours.

\begin{equation}
\label{eq:ConcaveVoidPhysics1}
S_b \buildrel \Delta \over = \{(b',\angle(b'~b~b^0)) : b' \in nbr(b)\}
\end{equation}

The mid-point $D_{pos}(b)$ is then calculated as below in Equation~\ref{eq:ConcaveVoidPhysics3}.

\begin{equation}
\label{eq:ConcaveVoidPhysics3}
D_{pos}(b) = \frac{1}{2}\mathlarger{\sum_{b' \in G_b}}{b'} 
\end{equation}

$D(b)$~is the \textit{concave reduction} vector is derived from the identified mid-point of the neighbour gap~($G_b$) to the parent agent~($b$) as shown in Equation~\ref{eq:ConcaveVoidPhysics3a}.

\begin{equation}\label{eq:ConcaveVoidPhysics3a}
D(b) = bD_{pos}(b) 
\end{equation}

The final vector needs to accommodate an agent's surroundings. Due to agent proximity it can be assumed that there are no agents in the path between the agent and the gap, however, there may be an obstacle in the agent's path. The \textit{concave-reduction-vector} must include an obstacle avoidance component~$v_o(b)$. As with all the previous vector based calculations a weighting is applied $k_{cr}$ to allow the model to vary the intensity of the effects. The resultant vector is normalised to produce the \textit{unit-movement-vector} as shown in~Equation~\ref{eq:ConcaveVoidPhysics4}.

>>>>

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{relsize}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{array}
%\usepackage{caption}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage{fixltx2e}
\usepackage{float}
\usepackage{url}
%\usepackage{becs}
\usepackage{listings}
\lstset{%
  basicstyle=\normalsize,
  numbers=left,
  numberstyle=\normalsize,
  stepnumber=1,
  xleftmargin=2em
}
%\usepackage{color}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
%\let\latexvec=\vec
%\let\vec=\latexvec
\interdisplaylinepenalty=2500
\newtheorem{definition}{Definition}
\newtheorem{exampleb}{Example}
\newcommand{\subsubsubsection}[1]{\medskip\par\textit{#1:}}
\newcommand{\card}[1]{\vert{#1}\vert}
\newcommand{\magn}[1]{\vert{#1}\vert}
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{xcolor}
\usepackage{standalone}

\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Concave reduction in Self-Healing Swarms}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, Michael Brockway\authorrefmark{1}, Alun Moon\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, Newcastle upon Tyne, NE1 8ST}
\markboth
{Eliot \headeretal: Concave reduction in Self-Healing Swarms}
{Eliot \headeretal: Concave reduction in Self-Healing Swarms}
\corresp{Corresponding author: Neil Eliot (e-mail: neil.eliot@northumbria.ac.uk).}

\begin{abstract}
%\boldmath
Swarms consist of many agents that interact using a simple set of rules that create emergent behaviours. One such behaviour makes a swarm tolerant of failures. Failures occur for many reasons, for example; resource exhaustion, component failure, or a disruption from an external event. The loss of agents reduces the size of a swarm and may create an irregular structure. A swarm's structure can also be irregular due to initial conditions, loss of agents or due to an obstacle. These changes in the structure or size do not stop a swarm from functioning but can adversely affect the efficacy. This paper introduces and demonstrates emergent behaviour that can counter the effect of agent loss or irregular structures. The behaviour causes concave deformities to be removed. This technique is implemented by each agent identifying its position in relation to its neighbours. If the agent meets a defined criterion, based on the local geometry of the agent's neighbours to decide when an anomaly is present, the agent's `normal' movement is altered in such a way as to remove a gap between the neighbours. These changes in direction of agents cause the swarm, as a whole, to be reorganised into a more uniform structure with no internal voids.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.
% Note that keywords are not normally used for peer review papers.
\begin{IEEEkeywords}
Swarming, Swarm Dynamics, Mobile Sensor Networks, Algorithms.
\end{IEEEkeywords}

\titlepgskip=-15pt
\maketitle

\section{Introduction}\label{sec:ConcaveReduction}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively.
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two \textit{field effects} to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. \textit{Field effects} have ranges around an agent to determine the effect other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}) the effects are derived from these ranges as vectors. All ranges must fall within the sensing capabilities of the agent ($S_b$). It is usual for the cohesion field to have a radius ($C_b$) which is larger than the repulsion radius ($R_b$). When an agent ($b'$) moves into the \textit{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has its repulsion vector applied and the agent will tend towards moving away from the obstacle.
\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}
\begin{figure}
%\begin{tikzpicture}
%	\node at (0,0) [circle,fill=black!50]{};
%\end{tikzpicture}
	\includegraphics{stableswarm}
\end{figure}
When cohesion and repulsion (\textit{inter-agent vectors}) are the only field effects creating a swarming effect the number of stable structures that can develop is limited. These structures, which are constructed from triangles, are either straight edges or partial lattices. Partial lattices can create concave anomalies (\textit{dents}) and convex anomalies (\textit{peaks}) in a perimeter~(Fig.~\ref{concave:SwarmStableShape1}). An anomaly is any construct within the swarm that is not a hexagonal lattice or an agent distribution that causes the swarm to deviate from having only convex or flat edges. The advantage of creating a well structured swarm can be realised in two classes of swarm application; \textit{Reconnaissance}, by creating a more uniform structure that eliminates blank spots, and \textit{Object Containment}, when using a swarm to surround a target.

@@

\section{Swarm Model}\label{sec:Swarm Model}
To describe the removal of voids a swarming model needs to be defined.

An agent $b$ in the swarm ($b\in\mathcal S$) has a sensor ($S_b$) that detects other agents
($b'$)in the swarm
determining their range $r$  and direction $\beta$.  From this data the agent generates a
set of neighbours ($\mathcal N_b$) that are within a specific range.  Usually the range of the
cohesion field $C_b$.

The Agent creates a set of neighbours containing range and bearing pairs
$(r,\beta)$ as shown in equation
\ref{set:Nb}
\begin{equation}
\mathcal N_b = \{ (r,\beta) \ldots \}
\label{set:Nb}
\end{equation}
These range and bearing pairs are the relative position vector for each
neighbour $b'$ with respect to the sensor reference frame of agent $b$.
Alternative notations are given in equation \ref{notation}
\begin{equation}
	\mathcal N_b = \{(r,\beta) \ldots \} = \{ (r_{b'}, \beta_{b'}) \ldots \} = \{
	\vec{b'} \ldots \}
	\label{notation}
\end{equation}

Equation~\ref{eq:BotPhysics1} defines a weighted model that includes cohesion,
repulsion, direction and obstacle avoidance ($\vec v_c(b), \vec v_r(b),
\vec v_d(b), \vec v_o(b)$). 
The weightings $k_c, k_r, k_d, k_o$ allow each component to be scaled to tailor the swarming effect.

\begin{equation}\label{eq:BotPhysics1}
  \vec v(b) = k_c\vec v_c(b) + k_r\vec v_r(b) + k_d\vec v_d(b) + k_o\vec v_o(b)
\end{equation}

\textit{Repulsion} $\vec v_r(b)$, defined in Equation~\ref{eq:Repulsion1}, is the
directional movement required to prevent agents colliding. $\mathcal R_b$ is
the set of agents that are within an agent's $b$ repulsion range.

\begin{equation}\label{eq:Repulsion1}
\vec v_r(b) = 
\frac{1}{\card{\mathcal R_b}}
\left(
	\mathlarger{\mathlarger{\sum_{\vec b' \in \mathcal R_b}}}
	{\left( 1-\frac{r_{b'}}{R_b} \right)}
	\vec b'
\right)
\end{equation}

\textit{Cohesion} $\vec v_{c}(b)$, defined in Equation~\ref{eq:FlyToCentre1},
calculates the movement that makes an agent move towards other agents to form
a cohesive structure. $\mathcal C_b$ is the set of agents that are within an agent's ($b$) cohesion range.

\begin{equation}\label{eq:FlyToCentre1}
	\vec v_c(b) =
	\frac{1}{\card{\mathcal C_b}}\left({\mathlarger{\sum_{\vec b' \in
	\mathcal C_b}}{\vec b'}}\right)
\end{equation}

\textit{Direction} $\vec v_d(b)$, defined in Equation~\ref{eq:Direction},
generates a directional vector for an agent to move towards a destination
defined as $\vec d$.

\begin{equation}
\label{eq:Direction}
\vec v_d(b) = \vec d
\end{equation}

@@

Equation~\ref{eq:Obstacle2} shows the repulsion vector $\vec v_o(b)$ for an agent.
$\mathcal O_b$ is the set of obstacles that are within the range of agent $b$.
The obstacles are identified using the distance between an agent and an
obstacle $o_r$ and comparing the result to the fixed obstacle repulsion
field~$O_b$, so $\forall \vec o \in \mathcal O_b : \magn{\vec o}\leq O_b$.
The applied repulsion is calculated by scaling the normalised sum of the
normalised vectors $\hat o$ by $O_b$.  Note that $\string^$ is the equivalent
of $\hat v = \frac{\vec v}{\magn{\vec v}}$ the normalised vector.

@@

\begin{eqnarray}\label{eq:Obstacle2}
  \vec v_o(b) & = & O_b \hat q_o \\
	\mathrm{where~}  \vec q_o & = & \sum_{\vec o\in \mathcal O_b } \hat o
	\nonumber \\
	\vec v_o(b) & = & O_b \widehat{\sum_{\vec o\in \mathcal O_b }} \hat o \nonumber
\end{eqnarray}

An agent's \textit{movement vector} is the sum of all the component vectors as
shown in Equation~\ref{eq:BotPhysics1}. This is a similar technique as used by
Hashimoto et. al.~\cite{HAY:08}. For a vector to be used for movement it be
normalised before the agent's speed ($s_b$)  can be applied. The resultant
\textit{movement vector} $\vec m_b$, as defined in Equation~\ref{eq:AgentMovement}, is calculated using unit time, speed and the normalised \textit{movement vector}.

\begin{equation}\label{eq:AgentMovement}
\vec m_b = s_b  \hat v(b)  t
\end{equation}

@@

The detection mechanism detects both the outer edge of a swarm (equation
\ref{perimeter-predicate})  and any internal features~(voids) that satisfy the same
set of conditions. It is therefore possible to have both voids and islands of
agents within the same swarm~(Fig.~\ref{fig:PerimeterBots2}). Voids are best
defined as perimeters that are concave (equation \ref{concave-predicate}) in nature and inside another perimeter. McLurkin describes two types of perimeters convex and concave~\cite{MD:09}. A convex perimeter is an edge where the average angle of the agent's exposed faces are~$> 180^\circ$. A concave perimeter is where the average exposed angle is~$< 180^\circ$.

@@

The set of neighbours $\mathcal N_b$ (eqn \ref{set:Nb}) is sorted into a
sequence $\mathcal P_a$ with the bearing values
increasing
\begin{equation}
	\mathcal P_a = \langle (r_0,\beta_0),\ldots,(r_n,\beta_n) \rangle
\end{equation}
such that $\forall i \in \{0\ldots n-1\}: \beta_i<\beta_{i+1}$

This set of agents forms the perimeter of an enclosing polygon of agent $b$. 
Each consecutive pair of agents (in the sequence) defines an edge, which has
length $d$ and an angle $\theta$ given by the difference in bearings of
succesive neighbours.  The sequence of edges
that forms this polygon is
\begin{equation}
	\mathcal{P}_e = \langle (d_0,\theta_0), \ldots , (d_n,\theta_n) \rangle
\end{equation}
where
\begin{equation}
	\theta_i = \beta_{i+1} - \beta_i
\end{equation}
the index addition is modulo $|\mathcal N_b|$ making $\beta_0$ the successor
bearing to $\beta_n$ ($n+1 = 0$).  The angles $\theta$ must lie in the range
$0<\theta\leq2\pi$.
This restriction on the values of $\theta$ enforce a condition that
\begin{equation}
	\sum\theta_i = 2\pi
\end{equation}
The length of a perimeter edge is given by the cosine rule
\begin{equation}
	d_i^2 = r_{i+1}^2 + r_i^2 -2r_{i+1}r_i \cos\theta_i
\end{equation}

An agent is on the perimeter of the swarm if it is not enclosed by the polygon
defined in $\mathcal{P}_e$.  Simple geometry shows that this is the case given
by the predicate in equation \ref{enclosing}
\begin{equation}
	\exists \theta_i \in \mathcal{P}_e : \theta_i\geq\pi
	\label{enclosing}
\end{equation}
The polygon is considered `open' if two succesive agents on the perimeter 
cannot see each onther, their separation $d$ is greater that the range of the
attractive field.  An open polygon does not enclose the agent $b$ so it
considered to be on the perimeter.

The agent $b$ is on the perimeter of the swarm if the predicate in equation
\ref{perimeter-predicate} is true.
\begin{equation}
	\exists d_i\in\mathcal{P}_e:d_i>C_b \vee
	\exists\theta\-i\in\mathcal{P}_e:\theta_i\geq\pi
	\label{perimeter-predicate}
\end{equation}
An agent is at the apex of a concave region of the perimeter if
\begin{equation}
	\exists(\theta_i,d_i)\in\mathcal{P}_e : d_i>C_b\wedge\theta_i<\pi
	\label{concave-predicate}
\end{equation}

\textit{Orientation independent}  If the agent $b$ is rotated through an angle of $\gamma$ then 
the bearings are rotated by $-\gamma$, \[ \beta_i\mapsto\beta_i-\gamma \]
The  angle between successive agents is now
\[
	\theta_i  =  (\beta_{i+1}-\gamma) - (\beta_i-\gamma)
	 = \beta_{i+1}-\beta_i-\gamma+\gamma
	 = \beta_{i+1}-\beta_i
\]
The predicates are independent of the orientation of the sensing agent.

@@

@@

When an agent is identified as being a component of a concave
characteristic~(eqn.~\ref{concave-predicate}) the normal \textit{movement-direction vector} is replaced by a \textit{concave reduction vector}. This new vector causes the agent to move in a direction that will reduce or remove a concave edge by moving the agent towards the identified gap. The effect of this will be to either straighten an outer perimeter or reduce/remove a void. This change in direction effects the distance and magnitude variances due to the changes induced. Figures~\ref{fig:InterAgentEffect1} and \ref{fig:InterAgentEffect2} shows this effect in more detail. Figure~\ref{fig:InterAgentEffect1}~shows the initial positions of the agents before the concave reduction is applied to the agent. Figure~\ref{fig:InterAgentEffect2}~shows the effects on the relationship with the agent's neighbours. The aggregate change is an increase in the inter-agent distances and an increase in the resultant magnitude effects.

@@

\section{Concave reduction mathematical model}\label{concave:ConcaveVoidReduction1}
% As part of identifying a perimeter each agent's neighbours must be sorted by angle. A pair of agents are identified as a set $b.gap$ ($b.gap \equiv G_b$). 
When an agent is on the perimeter, equation \ref{concave-predicate} is
satisifed, it is subject to concave reduction.
The pair of neighbours that form the concave element is $\mathcal G_b$, the
gap in the perimeter.
Equation~\ref{eq:ConcaveVoidPhysics3} calculates the centroid of the gap agents.

% Equation~\ref{eq:ConcaveVoidPhysics1} generates the dictionary set of all the neighbours along with an angle that each of the agent's neighbours make with the first detected neighbour. $nbr(b)$ returns a set consisting of an agent's ($b$) neighbours.

%\begin{equation}
%\label{eq:ConcaveVoidPhysics1}
%S_b \buildrel \Delta \over = \{(b',\angle(b'~b~b^0)) : b' \in nbr(b)\}
%\end{equation}

The mid-point $\vec D(b)$ is then calculated as below in Equation~\ref{eq:ConcaveVoidPhysics3}.

\begin{equation}
\label{eq:ConcaveVoidPhysics3}
\vec D(b) = \frac{1}{2}\sum_{\vec b' \in \mathcal G_b}\vec b' 
\end{equation}

$\vec D(b)$~is the \textit{concave reduction} vector is derived from the identified
mid-point of the neighbour gap~($\mathcal{G}_b$) to the parent agent~($b$) as
shown in Equation~\ref{eq:ConcaveVoidPhysics3}.

%%\begin{equation}\label{eq:ConcaveVoidPhysics3a}
%%D(b) = bD_{pos}(b) 
%%\end{equation}
%%\fbox{\textit{do we mean product of $b$ and $D_{pos}$?}}
